### 2023-01-25 paper-thesis/chapters/10-coc.tex
Work from yesterday

Finally, we define \verb|expectBetaEq|. This rule first beta reduces the heads of both sides, and then compares them. If the head is not the same, the rule fails. Otherwise, it recurses on the sub-expressions. One special case is when comparing two \verb|FnConstruct|s. Here we need to take into account alpha equality: two expressions which only differ in the names that they use should be considered equal. We implement this by substituting in the body of the functions, replacing their argument names with placeholders.

@@

>>>>

@@

\section{Beta equality}
\label{sec:coc-betaeq}

Finally, we define \verb|expectBetaEq|. This rule first beta reduces the heads of both sides, and then compares them. If the head is not the same, the rule fails. Otherwise, it recurses on the sub-expressions. One special case is when comparing two \verb|FnConstruct|s. Here we need to take into account alpha equality: two expressions which only differ in the names that they use should be considered equal. We implement this by substituting in the body of the functions, replacing their argument names with a unique placeholder.

This substitution is called \verb|AlphaEqVars : ID * ID -> Expr|. The combination of the ID on the left and right hand side of the equality guarantees that the substitution is unique. In figure \ref{fig:beta-eq-rules} we show how \verb|AlphaEqVars| is used to determine equality of functions.

\begin{figure}[h]
	\figuresection[\fbox{$\beq{\scope{s_1}{e_1}}{\scope{s_2}{e_2}}$}]{Beta equality rules}

	\begin{mathpar}
		\inferrule{
		} {
			\beq
			{ \co{AlphaEqVars}(x_1, x_2) }
			{ \co{AlphaEqVars}(x_1, x_2) }
		}

		\inferrule{
			\beq{\scope{s_1}{a_1}}{\scope{s_2}{a_2}}
			\\\\ s_1' = \co{sPutSubst}(s_1, x_1, (\co{sEmpty}(), \co{AlphaEqVars}(x_1, x_2)))
			\\\\ s_2' = \co{sPutSubst}(s_2, x_2, (\co{sEmpty}(), \co{AlphaEqVars}(x_1, x_2)))
			\\\\ \beq{
				\scope{s_1'}{b_1}
				}{ 
				\scope{s_2'}{b_2}
			}
		} {
			\beq
			{ \scope{s_1}{\co{FnType}(x_1, a_1, b_1)} }
			{ \scope{s_2}{\co{FnType}(x_1, a_2, b_2)} }
		}

		\inferrule{
			\beq{\scope{s_1}{a_1}}{\scope{s_2}{a_2}}
			\\\\ s_1' = \co{sPutSubst}(s_1, x_1, (\co{sEmpty}(), \co{AlphaEqVars}(x_1, x_2)))
			\\\\ s_2' = \co{sPutSubst}(s_2, x_2, (\co{sEmpty}(), \co{AlphaEqVars}(x_1, x_2)))
			\\\\ \beq{
				\scope{s_1'}{b_1}
			}{ 
				\scope{s_2'}{b_2}
			}
		} {
			\beq
			{ \scope{s_1}{\co{FnConstruct}(x_1, a_1, b_1)} }
			{ \scope{s_2}{\co{FnConstruct}(x_1, a_2, b_2)} }
		}
	\end{mathpar}

	\caption{Rules for beta equality in the Calculus of Constructions}
	\label{fig:beta-eq-rules}
\end{figure}

### 2023-01-25 paper-thesis/chapters/21-adding-inference.tex
Work from yesterday

\todo{higher-order pattern unification / miller pattern unification, what agda does}

@@

@@

Instead, we will be implementing a novel, less powerful form of first-order unification. This will work by explicitly denoting which variables \emph{could be} free, and explicitly handling these cases in a way that approximates first-order unification. We will denote this algorithm as \emph{approximated first-order unification (AFOU)}.

@@

>>>>

@@

A more powerful algorithm that could solve this is \emph{higher-order pattern unification} as defined by Miller~\cite{Miller89}, which is implemented in Agda. Implementing this in Statix is theoretically possible\footnote{since Statix is turing-complete}, but would be quite difficult to do in practice.

@@

Instead, we will be implementing a novel, less powerful form of first-order unification. This will work by explicitly denoting which variables \emph{could be} free, and explicitly handling these cases in a way that approximates first-order unification. We will name this algorithm \emph{approximated first-order unification (AFOU)}.

@@

\begin{lstlisting}
postulate f : (A : Type -> _: A -> _: A -> A);
postulate u : X : Bool -> Bool;
postulate v : Y : Bool -> Bool;
f _ u v
\end{lstlisting}

### 2023-01-25 paper-thesis/chapters/40-discussing-statix.tex
Meeting results
Work from yesterday

\chapter{The usability of Statix for defining dependently typed languages}

>>>>

\chapter{Ergonomics of Spoofax}

Now that we've implemented a feature-rich dependently typed language in Statix, this chapter discusses the experience.

\section{Statix}

- expressive language
- more control over metavariables
- nested constraints (datatypes reduce)
- generic functions? reverse/concat for lists
- debugging (statix tests are not great)

\section{Spoofax}

- errors
- performance
- testing works well
-
